---
title: "Dplyr"
author: "Davide Romano"
output: 
  html_document: 
    keep_md: yes
---

##DataCamp notes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load `dplyr` package and dataset `hflights`.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(hflights)
```

Convert data.frame into a tbl.

```{r}
hflights <- tbl_df(hflights)
hflights
```

`Glimpse()` on tbl is similar to `str()`.

```{r}
glimpse(hflights)
```

The `dplyr` package contains five key data manipulation functions:

* `select()` return a subset of the columns
* `filter()` return a subset of the rows
* `arrange()` reorders the rows according to single or multiple variables
* `mutate()` add columns from existing data
* `summarise()` reduces each group to a single row by calculating aggregate 

###`select()`

Example of `select()` function.

```{r}
select(hflights,ActualElapsedTime, AirTime, ArrDelay, DepDelay)
```

Helper functions that can help to select groups:

* `starts_with("X")`
* `ends_with("X")`
* `contains("X")`
* `matches("X")`
* `num_range("x", 1:5)`
* `one_of(x)`

###`mutate()`

Example of `mutate()` function.

```{r}
mutate(hflights, ActualGroundTime = ActualElapsedTime - AirTime)
```

###`filter()`

Example of `filter()` function.

```{r}
filter(hflights,Distance > 1000)
```

###`arrange()`

Example of `arrange()` function.

```{r}
head(arrange(hflights,DepDelay))
```

