---
title: "Dplyr"
author: "Davide Romano"
output: 
  html_document: 
    keep_md: yes
---

#DataCamp notes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load `dplyr` package and dataset `hflights`.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(hflights)
```

Convert data.frame into a tbl.

```{r}
hflights <- tbl_df(hflights)
hflights
```

`Glimpse()` on tbl is similar to `str()`.

```{r}
glimpse(hflights)
```

The `dplyr` package contains five key data manipulation functions:

* `select()` return a subset of the columns
* `filter()` return a subset of the rows
* `arrange()` reorders the rows according to single or multiple variables
* `mutate()` add columns from existing data
* `summarise()` reduces each group to a single row by calculating aggregate 

###`select()`

Example of `select()` function.

```{r}
select(hflights,ActualElapsedTime, AirTime, ArrDelay, DepDelay)
```

Helper functions that can help to select groups:

* `starts_with("X")`
* `ends_with("X")`
* `contains("X")`
* `matches("X")`
* `num_range("x", 1:5)`
* `one_of(x)`

###`mutate()`

Example of `mutate()` function.

```{r}
mutate(hflights, ActualGroundTime = ActualElapsedTime - AirTime)
```

###`filter()`

Example of `filter()` function.

```{r}
filter(hflights,Distance > 1000)
```

###`arrange()`

Example of `arrange()` function.

```{r}
head(arrange(hflights,DepDelay))
```

###`summarise()`

Example of `summarise()` function.

```{r}
summarise(hflights, min_dist = min(Distance),max_dist = max(Distance))
```

R contains many aggregating functions, as `dplyr` calls them:

* `min(x)` - minimum value of vector x.
* `max(x)` - maximum value of vector x.
* `mean(x)` - mean value of vector x.
* `median(x)` - median value of vector x.
* `quantile(x, p)` - pth quantile of vector x.
* `sd(x)` - standard deviation of vector x.
* `var(x)` - variance of vector x.
* `IQR(x)` - Inter Quartile Range (IQR) of vector x.
* `diff(range(x))` - total range of vector x.

An example to summarise the longest `Distance` for `diverted` flights.

```{r}
summarise(filter(hflights, Diverted==1), max_div=max(Distance))
```

`dplyr` provides several helpful aggregate functions:

* `first(x)` - The first element of vector x.
* `last(x)` - The last element of vector x.
* `nth(x, n)` - The nth element of vector x.
* `n()` - The number of rows in the data.frame or group of observations that summarise() describes.
* `n_distinct(x)` - The number of unique values in vector x.
* `sum()`
* `mean()`

Example:

```{r}
summarise(hflights, n_obs = n(), n_carrier = n_distinct(UniqueCarrier), n_dest = n_distinct(Dest))
```

<br>

##Pipes `%>%`

The following two commands that are completely equivalent:

```{r, message=FALSE, warning=FALSE}
mean(c(1, 2, 3, NA), na.rm = TRUE)
c(1, 2, 3, NA) %>% mean(na.rm = TRUE)
```

<br>

##`group_by()`

`group_by()` lets you define groups within your data set.

```{r}
hflights %>%
  group_by(UniqueCarrier) %>%
  summarise(p_canc = mean(Cancelled == 1) * 100) %>%
  arrange(p_canc)
```
